\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2016-04-26 doc/sw-cfhtls-referee-response.txt

We thank the referee for providing useful comments on the manuscript. We have implemented all of the suggestions and the textual changes are shown in bold in the revised version of the manuscript. Below are our answers to referee's comments.

-------------
In the abstract, I would replace "albeit at the cost of a higher false positive rates" by "albeit at the cost of a lower purity", the purity is a quantity well known by the astronomers and I think it makes things clearer here (especially for an abstract).

Ans: Done

The short Section 4 should become the 4th subsection of Section 3 with slight changes in the titles, like :
          3. Training Sample
                  3.1.  Methodology to simulate lenses
                  3.2. Simulated lens sample and catalog description [no change]
                  3.3. Limitation of the simulated lens sample [no change]
                  3.4. Duds and impostors
and by adding a word about duds at the end of the introduction of Section 3.

Ans: Done. This means that Sections 5,6,7 and 8 now become 4,5,6 and 7, respectively.

There is sometimes a confusion between real lensing galaxies and real galaxies chosen to be lensing galaxies in the training set of simulations. For example, in the last paragraph before subsubsection 3.1.2., in the sentence "the simulated image is added to the real CFHTLS image centered on the lensing galaxy", the authors should write 'lensing galaxy' between quotes or rephrase it : "..centered on the galaxy chosen to act as a lens". The same remarks applies to other occurrences in the same paragraph (and maybe others in the rest of the text).

Ans: Done.

The caption of Figure 2 should be completed : "Einsten radius and total magnification distributions..."

Ans: Done.

Last paragraph before subsection 3.3, first sentence, I would precise : "We produce catalogs with lens and source properties for each of the three types of SIMULATED lenses."

Ans: Done.

There is sometimes a confusion between the (evolving) posterior probability after each classification and the final posterior probability after stage 2, and both are noted P (even if I understand that somehow this is the same quantity). Maybe the authors could note the final posterior Pf and/or be more specific in the text. For example, in the caption of Figure 4, or, in subsection 5, paragraph 3, line 5 : "Each subject, after Stage 2, has a FINAL posterior probability P".

Ans: Done.

In Figure 4, there should be a better color contrast between the new candidates and the known candidates to make the results more readable.

Ans:  The low color contrast was chosen on purpose to suggest that these are all Space Warps identified candidates and the slight difference is to point out that one of them is a new sample and the other is a known one. The figure is modified to increase the color contrast but both are still shades of blue.

Page 9, subsection 5, 4 lines before the end of paragraph 2, I would precise "In stage 1, ..." before "...those images which cross these threshold values are 'retired'"

Ans: Done.

Maybe the authors should re-explain (Paper I) quickly at this stage of the article (3rd paragraph of subsection 5) that the stage 1 favors completeness while the stage 2 favors purity of the final sample.

Ans: Done.

Subsection 6.2, paragraph 3, penultimate sentence, I would precise : "...the atypical color and structure of the LENS galaxy"

Ans: Done.

At the end of subsection 6.4., as well as in the conclusion, the authors quickly deal with the subject of the deflector subtraction. I would like them to make clearer in their article that this is not a trivial issue in the detection of galaxy-scale lenses, and that solutions are being actively sought for a better subtraction of this foreground light.

Ans: A sentence is added, as per the suggestion, towards the end of subsection 5.4 (which was 6.4 earlier). The conclusion part is left unchanged as the galaxy subtraction is mentioned only for completeness, to merely point out the main difference between Space Warps and Ringfinder.

There is sometimes a confusion between the various subsamples considered in the article. For example, in Subsection 6.5, paragraph 2, lines 4-5, it is hard to distinguish in the text if the "known" samples considered refer to confirmed lenses or the whole candidates.

Ans: Done.

In subsection 6.5, describing Figure 7, could the authors say one word about the few points at lower image separation which are not consistent with the "total" profile?

Ans: Done.

There is some overlap between Section 6 and Section 7, even if the developments are interesting. But maybe Section 7 should be somehow shortened for clarity?

Ans: We have removed the 2nd and 3rd paragraph from the Discussion, Sect. 6.1 (which was 7.1 earlier) to avoid repetition

The title of subsubsection 7.3.2. "Blind lens search" seems a bit too general and maybe not relevant. It should be replaced by something more specific like "Lens positions in the image cutouts".

Ans: Done.

In Figure 11 and its description in the text (subsubsection 7.3.3, paragraphs 4 and 5), the distinction between the red and green curves is unclear, especially for what concerns the detected lenses. There should be a word about it in the caption of the figure too.

Ans: The text in para 4 is modified to explain better the difference for the missed sample. The beginning and the end of para 5 state the difference between the green and the red for the detected sample. The caption of the figure is modified too.

Subsubsection 7.3.3, paragraph 7, I think there is a mistake in "i) The ratio of large positive kicks to small positive kicks for the detected sample is higher than in the missed sample"... I think the authors meant "negative" instead of the second "positive"?!

Ans: Text is corrected.

Last paragraph before Section 8, I would again replace "increase the rate of false positives" by "decrease the purity".

Ans: Done.

The authors could mention EUCLID with the other future surveys in their conclusion.

Ans: Done.

### 2016-04-26 doc/sw-system-referee-response.txt

We thank the referee for their very helpful comments, all of which we
address below.  Their comments are shown unindented, while our
responses are shown indented.

In addition, we clarified the purity calculation with the assumptions
we made about the expected number of lenses present in the survey, and
also made it clear that the purity of the Stage 1 sample (15%) was
chosen to be lower than the maximum purity obtainable. The abstract
and conclusions remain unchanged, as these refer to the final purity,
after Stage 2 refinement. We include a PDF version of the paper with
all changes highlighted, for the referee's convenience.

### 2016-04-26 doc/sw-cfhtls.tex

The bottom panels of \Fref{fig:detmis}, show the trajectories of missed
lenses for the cases which are visually easier (red) and more
difficult (green) to identify.  In spite of some mild
qualitative differences, both set of trajectories have very similar
behaviour. The trajectories in panel (e) are typical of this sample in
terms of Nclass and the dominance of small negative kicks. Panel (f)
represents a small fraction of this sample where the kicks are only
small and negative. The panel (g) shows how some lenses receive a bunch
of large positive kicks which are led to rejection by still mostly small
negative kicks. Finally, panel (h) shows those cases of lenses which
received almost sufficient number of large positive kicks to be detected
but ended up being rejected.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.6]{det_mis_traj.pdf}
\includegraphics[scale=0.6]{det_mis_hist.pdf}
\caption{ \label{fig:detmis}
Left: Examples of \StageOne trajectories of some known lenses.  Upper
panels show detected lenses, lower panels show missed lenses (false
negatives). The green trajectories for the detected sample are
counterparts of the red and green trajectories of the missed sample
except that the kicks are positive for the detected sample. The red
trajectories of the detected sample demonstrate that sufficient number
of positive large kicks can lead to the detection of the lenses. Within the
missed sample, the red and green correspond to visually easier and more
difficult to identify systems, respectively. Right: Histogram of classification
power received by known lenses.  Among the sample of missed lenses, most
classifications are from the low power volunteers identifying these
images incorrectly (i.e.  $-\Delta$~log(P)) which overshadows the small
number of correct identifications (i.e.  $\Delta$~log(P)) coming from
both the less and the high power volunteers combined. For the detected
lens sample, most classifications are correct identifications coming
from both low and high power volunteers. }
\end{center}
\end{figure*}

>>>>

The bottom panels of \Fref{fig:detmis}, show the trajectories of missed
lenses for the cases which are visually easier (light green) and more
difficult (dark green) to identify.  In spite of some mild
qualitative differences, both set of trajectories have very similar
behaviour. The trajectories in panel (e) are typical of this sample in
terms of Nclass and the dominance of small negative kicks. Panel (f)
represents a small fraction of this sample where the kicks are only
small and negative. The panel (g) shows how some lenses receive a bunch
of large positive kicks which are led to rejection by still mostly small
negative kicks. Finally, panel (h) shows those cases of lenses which
received almost sufficient number of large positive kicks to be detected
but ended up being rejected.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.6]{det_mis_traj.pdf}
\includegraphics[scale=0.6]{det_mis_hist.pdf}
\caption{ \label{fig:detmis}
Left: Examples of \StageOne trajectories of some known lenses.  Upper
panels show detected lenses, lower panels show missed lenses (false
negatives). The green trajectories for the detected sample are
counterparts of the light and dark green trajectories of the missed sample
except that the kicks are positive for the detected sample. The red
trajectories of the detected sample demonstrate that sufficient number
of positive large kicks can lead to the detection of the lenses. Within the
missed sample, the light and dark green correspond to visually easier and more
difficult to identify systems, respectively. Right: Histogram of classification
power received by known lenses.  Among the sample of missed lenses, most
classifications are from the low power volunteers identifying these
images incorrectly (i.e.  $-\Delta$~log(P)) which overshadows the small
number of correct identifications (i.e.  $\Delta$~log(P)) coming from
both the less and the high power volunteers combined. For the detected
lens sample, most classifications are correct identifications coming
from both low and high power volunteers. }
\end{center}
\end{figure*}

### 2016-04-26 doc/sw-system.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{CFHTLS_2013-10-29_101635_trajectories.pdf}
\caption{Typical \SW \StageOne subject trajectories. Top: 200 randomly-selected
subjects drift downwards as they are classified, while being nudged left and
right in probability by the agents as they interpret the volunteers input. The
dotted vertical lines show (left to right) the retirement threshold, prior
probability, and the detection threshold. Different colours denote the different
kinds of subject. Bottom panel: histograms of all the subject probabilities
computed to date, sub-divided by subject kind.}
\label{fig:swap:subject-trajectories}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage1_ROC.pdf}
\end{minipage}\hfill
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage2_ROC.pdf}
\end{minipage}
\caption{Receiver operating characteristic curves for the \SW system, using
the \cfhtls training set. Left: \StageOne, right: \StageTwo. Insets show a
magnified view of the top lefthand corner of each plot. Linestyles illustrate
the software agent properties (initial confusion matrix elements, learning or not)
discussed in the text, as well as the difference between the online and offline
analyses. The \StageTwo sample was defined using the online \StageOne results
(solid blue curve), while the offline analysis was chosen for the final
\StageTwo results (dot-dashed orange curve).
Lens probability threshold varies along the curves: the stars indicate the point where
the value of this quantity is 0.95.}
\label{fig:results:sample:roc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We first quantify the performance of the \SW system in terms of the recovery of
the training set images. At \StageOne, this set contained around 5712 simulated
lenses, and 450 duds; at \StageTwo, we used 152 images of simulated lenses and
201 duds selected as \StageOne false positives (\Sref{sec:design:stages:two}).
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for \cfhtls \StageOne and \StageTwo. These plots show the true positive
rate (TPR, the number of sims correctly detected divided by the total number of
sims in the training set), and the false positive rate (FPR, the number of duds
incorrectly detected divided by the total number of duds in the training set), both
for a given sample of detections defined by a particular probability threshold,
which varies along the curves.  In both stages, these curves show that true
positive rates of  around 90\% were achieved, at very low false positive rates.
The probability threshold that was used for retirement during the \StageOne online analysis
was 0.95; this point is marked with a star on each curve.
This turned out to be close to optimal (although a better approach would be to
keep track of the ROC curve as the survey progressed!).

@@

We thank Stuart Lynn, Kelly Borden, Laura Whyte, Brooke Simmons, David Hogg,
Daniel Foreman-Mackey, Thomas Jennings, Layne  Wright, Cecile Faure, Jonathan
Coles, Stuart Lowe and Jean-Paul Kneib for many useful conversations about
citizen science and gravitational lens detection, and to the Dark Energy Survey
and Pan-STARRS strong lensing science teams for their suggestions and
encouragement.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{CFHTLS_2013-10-29_101635_trajectories.pdf}
\caption{Typical \SW \StageOne subject trajectories. Top: 200 randomly-selected
subjects drift downwards as they are classified, while being nudged left and
right in probability by the agents as they interpret the volunteers input. The
dotted vertical lines show (left to right) the retirement threshold, prior
probability, and the detection threshold. Different colours denote the different
kinds of subject. Bottom panel: histograms of all the subject probabilities
computed to date, sub-divided by subject kind. The blue bar (of correctly-detected
sims) hides a grey bar of around 3000 new lens candidates, which are the
subject of \PaperTwo.}
\label{fig:swap:subject-trajectories}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage1_ROC.pdf}
\end{minipage}\hfill
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage2_ROC.pdf}
\end{minipage}
\caption{Receiver operating characteristic curves for the \SW system, using
the \cfhtls training set. Left: \StageOne, right: \StageTwo. Insets show a
magnified view of the top lefthand corner of each plot. Linestyles illustrate
the software agent properties (initial confusion matrix elements, learning or not)
discussed in the text, as well as the difference between the online and offline
analyses. The \StageTwo sample was defined using the online \StageOne results
(solid blue curve), while the offline analysis was chosen for the final
\StageTwo results (dot-dashed orange curve).
Lens probability threshold varies along the curves: the symbols
indicate the points where
the value of this quantity is 0.95. In the lefthand panel, the circle
point is masked by the square point.}
\label{fig:results:sample:roc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We first quantify the performance of the \SW system in terms of the recovery of
the training set images. At \StageOne, this set contained around 5712 simulated
lenses, and 450 duds; at \StageTwo, we used 152 images of simulated lenses and
201 duds selected as \StageOne false positives (\Sref{sec:design:stages:two}).
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for \cfhtls \StageOne and \StageTwo. These plots show the true positive
rate (TPR, the number of sims correctly detected divided by the total number of
sims in the training set), and the false positive rate (FPR, the number of duds
incorrectly detected divided by the total number of duds in the training set), both
for a given sample of detections defined by a particular probability threshold,
which varies along the curves.  In both stages, these curves show that true
positive rates of  around 90\% were achieved, at very low false positive rates.
The probability threshold that was used for retirement during the \StageOne online analysis
was 0.95; this point is marked with a symbol on each curve.
This turned out to be close to optimal (although a better approach would be to
keep track of the ROC curve as the survey progressed!).

@@

We thank Matthias Tecza, Stuart Lynn, Kelly Borden, Laura Whyte, Brooke Simmons, David Hogg,
Daniel Foreman-Mackey, Thomas Jennings, Layne Wright, Cecile Faure, Jonathan
Coles, Stuart Lowe, Alexander Fritz and Jean-Paul Kneib for many useful conversations about
citizen science and gravitational lens detection, and for helping guide the
discussion at \SW \Talk,
and to the Dark Energy Survey
and Pan-STARRS strong lensing science teams for their suggestions and
encouragement. We are grateful to the anonymous referee, whose comments
and suggestions helped improve the paper significantly.
